Quantitative processing of broadband data as implemented in a scientific splitbeam echosounder

### 2022-03-17 Paper/Manuscript.tex
Updates to manuscript, added figures

\author{Geir Pedersen}
\affiliation{Ecosystem Acoustics, Institute of Marine Research, Bergen, 5001, Norway}

@@

\author{Geir Pedersen}
\author{Rolf Korneliussen}
\author{Gavin J. Macaulay}
\author{Egil Ona}
\author{Nils Olav Handegard}
\affiliation{Ecosystem Acoustics, Institute of Marine Research, Bergen, 5001, Norway}

@@

@@

@@

@@

@@

@@

Compensation of spherical spreading loss requires compensation of received power by a factor of $r_c^2$, and hence compensation of amplitude by a factor of $\range_c$:
%
\begin{equation}
\label{eq:spreadcomp}
\ypcspread(\samplesymt) = \ypc(\samplesymt)\range_c(\samplesymt).
\end{equation}
%
where $\ypcspread(\samplesymt)$ is the pulse compressed signal compensated for spherical spreading. A discrete Fourier transform is performed on the range compensated pulse compressed sample data using a normalized sliding Hanning window, $\hannw(\genidxsym)$. The duration, $\tslide$, of the sliding window is chosen as a compromise between along-beam range resolution and frequency resolution. We suggest that it be at least twice the pulse duration and for computational efficiency reasons should result in a number of samples, $\nw$, which is a power of 2.

The normalised Hanning window, $\hannwnorm$, is given by: 
%
\begin{equation}
\label{eq:hannw}
\hannwnorm(\genidxsym) = \frac{\hannw(\genidxsym)}{\left( \frac{||\hannw||_2}{\sqrt{\nw}} \right)}, i = \frac{-\nw}{2}, \ldots, \frac{\nw}{2}
\end{equation}
%
and the discrete Fourier transform of the windowed data, $\ypcvolumef(\samplesymf)$, is then obtained from:
%
\begin{equation}
\label{eq:FFT_volume}
\ypcvolumef(\samplesymf) = \dft_\ndft 
\left( \hannwnorm(\genidxsym) \left(\ypcspread (\genidxsym+\samplesymt) \left[ u(\genidxsym + \frac{\nw}{2}) - u(\genidxsym - \frac{\nw}{2}) \right] \right) \right),
\end{equation}
%
where $u(\genidxsym)$ is the step function and $\samplesymt$ is the sample data index for the centre of the sliding window. The discrete Fourier transform of the auto correlation function of the matched filter signal, $\ymfautof(\samplesymf)$, also needs to be evaluated at the same frequencies:
%
\begin{equation}
\label{eq:FFT_TX_Auto}
\ymfautof(\samplesymf)  =  \dft_\ndft (\ymfauto(\samplesymt)).
\end{equation}
%
TODO: How to index samples along range? In the code \_n is used to indicate the range component of the FFt's. Needs attention.

\section{Illustrative examples}

A frequency modulated pulse scattered by a metallic sphere will exhibit frequencies at which very little energy is returned due to destructive interference \citep{stanton2008}. This is visible in the $\ts$ (Fig. \ref{fi:ts_example}a, \ref{fi:ts_example}b) and agrees well with theoretical estimates of the backscatter from spheres \citep{maclennan1981}. The amplitude of the backscatter signal also clearly shows these nulls (Fig. \ref{fi:ts_example}c), which are readily visible here due to the use of a linear chirp where time through the pulse corresponds to specific frequencies. The marked increase in range resolution is apparent once pulse compression has been applied (Fig. \ref{fi:ts_example}d), as are the temporal effects of the pulse compression operation.

A metallic sphere is a rather simple and ideal scatterer and we also present $\sv$ from a school of Atlantic mackerel (\textit{Scomber scombrus}) (Fig. \ref{fi:sv_example}a). The trend for increasing $\sv$ with frequency is well-known \citep{korneliussen2010} and is consistent with the trend observed in this example. In contrast to data from isolated scatterers, such as metallic spheres, the benefit of pulse compression on the backscatter from an object that generates many overlapping echoes is not immediately obvious (Fig. \ref{fi:sv_example}b, \ref{fi:sv_example}c), although in regions where the fish density decreases (e.g., top left of the school), single target echoes become visible.

\begin{figure}
\includegraphics[width=16cm]{Fig_ts_example}
\caption{\label{fi:ts_example}(color online) Target strength of 57.2 mm (a) and 22.0 mm (b) diameter tungsten carbide (with 6\% cobalt binder) calibration spheres, theoretical (solid line) and measured and derived using Eq. (\ref{eq:TS_f}) (dotted line). Echogram of non pulse compressed backscatter [dB re 1 W] (c) and pulse compressed backscatter [dB re 1 $\textrm{m}^2$] (d) from 57.2 mm (lower band) and 22.0 mm (upper band) diameter tungsten carbide spheres using a 2.048 ms duration linear frequency modulated pulse (160--260 kHz). The range from the spheres to the transducer varied during the time period shown.}
\end{figure}

\begin{figure}
%\includegraphics[width=16cm]{Figure5}
\caption{\label{fi:sv_example}(color online) (a): Mean frequency response (solid line) and standard error (dashed line) from a school of Atlantic mackerel (\textit{Scomber scombrus}) from three simultaneously operating linear frequency modulated echosounder channels (90--170, 160--260, and 280--450 kHz). Echogram of backscattered power [dB re 1 W] (b) and pulse compressed $\sv$ [dB re 1 $m^{-1}$] (c) from the 160--260 kHz pulse. The school is the trapezoid-shaped object between 25 and 90 m below the surface. The echo from the seafloor is visible at 130 m depth.} \end{figure}

>>>>

\author{Geir Pedersen}
\affiliation{Marine Ecosystem Acoustics, Institute of Marine Research, Bergen, 5001, Norway}

@@

\author{Rolf Korneliussen}
\author{Gavin J. Macaulay}
\author{Egil Ona}
\affiliation{Marine Ecosystem Acoustics, Institute of Marine Research, Bergen, 5001, Norway}

\author{Ruben Patel}
\affiliation{Codelab, Bergen, Norway}

\author{Nils Olav Handegard}
\affiliation{Marine Ecosystem Acoustics, Institute of Marine Research, Bergen, 5001, Norway}

@@

To illustrate each step of the process to obtain $TS(f)$ we use a single ping from a data file where a single metallic calibration sphere is suspended approximately 6 m below a 120 kHz transducer (fig.~\ref{Fig_TS_echogram}).

\begin{figure}
\includegraphics[width=16cm]{Fig_TS_echogram}
\caption{\label{Fig_TS_echogram} Sv as a function of frequency ($m$) and range ($n$) for a echogram raw data file. The red vertical line indicates ping to be processed in this section for TS(f). Single target calibration sphere is seen at approximately 6 m range and the seafloor at approximately 12 m.}
\end{figure}

@@

A frequency modulated pulse scattered by a metallic sphere will exhibit frequencies at which very little energy is returned due to destructive interference \citep{stanton2008}. This is visible in the $\ts$ (Fig. \ref{fi:SED}) and agrees well with theoretical estimates of the backscatter from spheres \citep{maclennan1981}. The amplitude of the backscatter signal also clearly shows these nulls (Fig. \ref{fi:ts_example}c), which are readily visible here due to the use of a linear chirp where time through the pulse corresponds to specific frequencies.

@@

As a metallic sphere is a rather simple and ideal scatterer, we present $\sv$ from a school of non-swimbladdered fish (\textit{tobis}) (fig.~\ref{Fig_Sv_echogram}) in the processes to obtain $Sv((f)$.

@@

\begin{figure}
\includegraphics[width=16cm]{Fig_Sv_echogram}
\caption{\label{Fig_Sv_echogram} Sv as a function of frequency ($m$) and range ($n$) for a echogram raw data file. The red vertical line indicates ping to be processed in this section.}
\end{figure}

@@

Compensation of spherical spreading loss requires compensation of received power by a factor of $r_c^2$, and hence compensation of amplitude by a factor of $\range_c$:
%
\begin{equation}
\label{eq:spreadcomp}
\ypcspread(\samplesymt) = \ypc(\samplesymt)\range_c(\samplesymt).
\end{equation}
%
where $\ypcspread(\samplesymt)$ is the pulse compressed signal compensated for spherical spreading. A discrete Fourier transform is performed on the range compensated pulse compressed sample data using a normalized sliding Hann window, $\hannw(\genidxsym)$. The duration, $\tslide$, of the sliding window is chosen as a compromise between along-beam range resolution and frequency resolution. We suggest that it be at least twice the pulse duration and for computational efficiency reasons should result in a number of samples, $\nw$, which is a power of 2.

The normalised Hann window, $\hannwnorm$, is given by: 
%
\begin{equation}
\label{eq:hannw}
\hannwnorm(\genidxsym) = \frac{\hannw(\genidxsym)}{\left( \frac{||\hannw||_2}{\sqrt{\nw}} \right)}, i = \frac{-\nw}{2}, \ldots, \frac{\nw}{2}
\end{equation}
%
and the discrete Fourier transform of the windowed data, $\ypcvolumef(\samplesymf)$, is then obtained from:
%
\begin{equation}
\label{eq:FFT_volume}
\ypcvolumef(\samplesymf) = \dft_\ndft 
\left( \hannwnorm(\genidxsym) \left(\ypcspread (\genidxsym+\samplesymt) \left[ u(\genidxsym + \frac{\nw}{2}) - u(\genidxsym - \frac{\nw}{2}) \right] \right) \right),
\end{equation}
%
where $u(\genidxsym)$ is the step function and $\samplesymt$ is the sample data index for the centre of the sliding window. The discrete Fourier transform of the auto correlation function of the matched filter signal, $\ymfautof(\samplesymf)$, also needs to be evaluated at the same frequencies:
%
\begin{equation}
\label{eq:FFT_TX_Auto}
\ymfautof(\samplesymf)  =  \dft_\ndft (\ymfauto(\samplesymt)).
\end{equation}

%
TODO: How to index samples along range? In the code \_n is used to indicate the range component of the FFt's. Needs attention.

By selecting a set of centre samples $n$, $Sv$ values can be presented as a function of range (n) and frequency (m) for each ping. The range for centre samples n could be chosen to half the window length or any other grid that the user prefer the data presented to be in. This can be useful when combining the Sv(f) across a range of transducers. In our example we have simply chosen n to be on the same order of the range resolution (fig.~\ref{Fig_Sv_m_n}).

\begin{figure}
\includegraphics[width=16cm]{Fig_Sv_m_n}
\caption{\label{Fig_Sv_m_n} Sv as a function of frequency ($m$) and range ($n$) for a single ping}
\end{figure}

For echo integration purposes it is common to integrate the Sv over a range, in our example we integrate the Sv from 15 to 34 m covering a school of non-swimbladdered fish (fig). We note that we normally require averaging over several individual pings to obtain an unbiased estimate of Sv for echo integration purposes. Here we present the sv(f) over the layer for our example ping for illustrative purposes (fig.~\ref{Fig_Sv_avg}). Even though this is for a single ping, we still observe a positive slope of the frequency response indicative of non-swimbladdered fish.

\begin{figure}
\includegraphics[width=16cm]{Fig_Sv_avg}
\caption{\label{Fig_Sv_avg} Sv as a function of frequency ($m$) averaged over a range interval covering a fish school }
\end{figure}

The trend for increasing $\sv$ with frequency is well-known \citep{korneliussen2010} and is consistent with the trend observed in this example. In contrast to data from isolated scatterers, such as metallic spheres, the benefit of pulse compression on the backscatter from an object that generates many overlapping echoes is not immediately obvious (fig.~\ref{Fig_Sv_echogram}).

In this implementation (ref code) several choices have been made. Some of these are well founded in signal processing literature whereas others are of a more practical and ad-hoc nature, where other approaches could been equally valid. These include, as examples,\\ 
Transmit pulse\\
Tapering of the transmit pulse\\
Decimation factors and filters\\
When Convoling the recieved signal with the transmit pulse - edge cases\\
Assuemd four sector transducer\\
Matched load - hvor stor effekt å gå utover \\
Assumed an impedance (trancuder)\\
Effective pulse duration...\\
Assuming SED algorithm, know the target depth. Improvements possible (SED for broadband). Number of samples for pulse comp - what is included, dense aggregations, sidelobes - pc, ...\\
Window size\\
Edge cases - removed\\

@@

...
